Direct-style process creation on Linux

### 2020-05-07 research/paper/paper.tex
notes and paper

Fork-style (such as \texttt{fork}) and spawn-style (such as \texttt{posix\_spawn}) each have advantages and disadvantages.
Correct and performant usage of fork-style places significant constraints on the calling process.
Spawn-style lifts many of these constraints,
but is fundamentally limited in the kinds of processes it can create.
With both fork-style and spawn-style,
it is difficult to express
branching behavior based on results or errors encountered while initializing the new process;
fork-style requires IPC if it wishes to communicate errors back to the rest of the program,
while robust indication of errors in spawn-style requires substantial growth in the spawn interface.

Direct-style is rare in general, but relatively common on academic operating systems.
Direct-style involves creating a new child process in a non-running state,
and then calling individual syscalls from the parent to initialize the child.
Direct-style does not constrain the calling process,
can create any kind of process supported by the underlying system,
and easily expresses errors in any individual step of the initialization.

Our contribution is an implementation of direct-style process creation for Linux.
Our implementation has all the advantages of direct-style
over Linux's fork-style \texttt{fork} syscall and spawn-style \texttt{posix\_spawn} libc function.

In our implementation, a parent process calls the \texttt{clone} syscall
to create a child process which shares memory and other resources with the parent,
and which runs no user code.
The parent then uses \texttt{rsyscall}, a cross-process syscall library,
to call arbitrary syscalls in the child process to initialize it;
for example, the parent might call \texttt{chdir} to change the child's current working directory.
After the process is fully initialized,
the parent can call \texttt{execve} in the child,
causing the process to start running an arbitrary executable
and detaching the child from the control of \texttt{rsyscall}.

We have used direct-style process creation on Linux to do TODO and TODO,
and we've benchmarked our implementation as TODO.

Our implementation is immediately deployable on recent Linux kernels,
and is distributed as a feature of the open source \texttt{rsyscall} library,
available from \url{https://github.com/catern/rsyscall}.

This paper is organized in several sections.
In section \ref{background}, we provide background on fork-style, spawn-style, and direct-style process creation,
and consider their advantages and disadvantages in more depth.
In section \ref{examples} we show examples of usage of direct-style process creation on Linux,
including use of process features which are difficult to use on Linux with \texttt{fork} or \texttt{posix\_spawn},
and briefly examine the user API.
In section \ref{implementation}, we explore our implementation in depth,
and discuss several difficult process-creation related details of \texttt{rsyscall}'s implementation.
In section \ref{evaluation}, we evaluate, somehow (TODO).
In section \ref{future_work},
we list several directions for future work related to direct-style process creation and to \texttt{rsyscall}.
In section \ref{conclusion}, we briefly and optimistically conclude.

@@

\section{Overview and examples}\label{examples}
In most operating system APIs, including the ones available on Linux,
when a program makes a syscall, it implicitly operates on the current process.
To perform direct-style process creation on Linux,
we need a Linux API where we instead explicitly specify in which process we want to make a syscall.

@@

\subsection{Example: Create and exec}
\begin{lstlisting}[float,language=Python,label={basic},caption={Creating a new process and immediately execing}]
# Use the clone syscall in the local thread to create a new thread;
# we use a wrapper that supplies defaults for all arguments.
child = local.thread.clone()
# Call execv to run a different executable in the child thread;
# We pass the executable path as the first argument in the argument list, as is traditional.
child.execv(hello_path, [hello_path])
\end{lstlisting}
In Listing \ref{basic}, we simply create a new process under our control (a thread)
and immediately exec a binary with it.
As exceptions are used for error-handling in the Python API,
there is no need for error-checking code.
%% **** note to self: concepts introduced in this example
%% local.thread
%% clone
%% execve
\subsection{Example: Passing down FDs}
\begin{lstlisting}[float,language=Python,label={fds},caption={Passing down FDs}]
child = local.thread.clone()
sock = child.socket(AF.INET, SOCK.DGRAM)
# Bind the socket to a sockaddr_in;
# the sockaddr is allocated in memory with child.ptr and is garbage collected.
sock.bind(child.ptr(SockaddrIn(0, 0)))
sock.listen(10)
sock.disable_cloexec()
child.execv(executable_path, [executable_path, "--listening-socket", str(int(sock))])
\end{lstlisting}
In Listing \ref{fds}, we create a new thread,
then create a listening socket bound to a random port in that thread,
then call exec, 
passing down the socket by disabling cloexec and passing its file descriptor number as an argument to the new program.

File descriptors, here, are object oriented and have relevant syscalls as methods.
They make syscalls in the process they are created in by default;
we can create more objects referring to the same file descriptor from different processes
if we want to make the syscalls from another process.
%% **** note to self: concepts introduced in this example
%% disable_cloexec
%% socket creation
%% using int on sock
%% .ptr
\subsection{Example: Pipeline}
\begin{lstlisting}[float,language=Python,label={pipe},caption={Creating a pipeline}]
# create the pipe
pipe = local.thread.pipe()
child1 = local.thread.clone()
# inherit the write-end of the pipe to child1, and replace child1.stdout with it
child1.inherit_fd(pipe.write).dup2(child1.stdout)
child1_proc = child1.execv(yes_path, [yes_path])
child2 = local.thread.clone()
# inherit the read-end of the pipe to child2, and replace child2.stdin with it
child2.inherit_fd(pipe.read).dup2(child2.stdin)
child2_proc = child2.execv(head_path, [head_path, "-n", "15"])
\end{lstlisting}
In Listing \ref{pipe},
we do the same as the Unix shell pipeline "yes | head -n 15".
We create a pipe,
then create two threads,
connect them with a pipe,
and exec a different program in each thread.

@@

We bind to a privileged port on localhost inside the namespace,
and create one child to listen on that socket,
and another child to connect to it.
%% **** note to self: concepts introduced in this example
%% network namespace
%% nested clone
% \subsection{Example: Starting a complex system}
% \begin{lstlisting}[float,language=Python,label={miredo},caption={Starting a complex system}]
% ### create socket outside network namespace that Miredo will use for internet access
% inet_sock = local.thread.socket(AF.INET, SOCK.DGRAM)
% inet_sock.bind(local.thread.ptr(SockaddrIn(0, 0)))
% # set some miscellaneous additional sockopts that Miredo wants
% set_miredo_sockopts(local.thread, inet_sock)
% ### create main network namespace thread
% ns_thread = local.thread.clone(CLONE.NEWNET|CLONE.NEWUSER)
% ### create in-network-namespace raw INET6 socket which Miredo will use to relay pings
% icmp6_fd = ns_thread.socket(AF.INET6, SOCK.RAW, IPPROTO.ICMPV6)
% ### create in-network-namespace socket,
% ### which Miredo will use for unassociated Ifreq ioctls
% reqsock = ns_thread.socket(AF.INET, SOCK.STREAM)
% ### create and set up the TUN interface
% tun_fd, tun_index = make_tun(ns_thread, "miredo", reqsock)
% ### create socketpair which Miredo will use to communicate
% ### between privileged process and Teredo client
% privproc_pair = ns_thread.socketpair(AF.UNIX, SOCK.STREAM)
% ### start up privileged process which manipulates the network setup in the namespace
% privproc_thread = ns_thread.clone()
% # We preserve NET_ADMIN capability over exec so that privproc can
% # manipulate the TUN interface. add_to_ambient_caps is a helper function
% # used because manipulating Linux ambient capabilities is fairly verbose
% add_to_ambient_caps(privproc_thread, {CAP.NET_ADMIN})
% # privproc expects to communicate with the main client over stdin and stdout
% privproc_side = privproc_thread.inherit_fd(privproc_pair.first)
% privproc_side.dup2(privproc_thread.stdin)
% privproc_side.dup2(privproc_thread.stdout)
% privproc_child = privproc_thread.execv(miredo_privproc_executable_path, [
%     miredo_privproc_executable_path, str(tun_index)
% ])
% ### start up Miredo client process which communicates
% ### over the internet to implement the tunnel
% # the client process doesn't need to be in the same network namespace,
% # since it is passed all the resources it needs as fds at startup.
% client_thread = ns_thread.clone(CLONE.NEWUSER|CLONE.NEWNET|CLONE.NEWNS|CLONE.NEWPID)
% # lightly sandbox by unmounting everything except for the executable
% # and its deps (known via package manager)
% unmount_everything_except(client_thread, miredo_exec.run_client.executable_path)
% # a helper function for preparing the fds that are passed as command line arguments
% async def pass_fd(fd: FileDescriptor) -> str:
%     client_thread.inherit_fd(fd).disable_cloexec()
%     return str(int(fd))
% client_child = client_thread.execv(miredo_client_executable_path, [
%     miredo_client_executable_path,
%     pass_fd(inet_sock), pass_fd(tun_fd), pass_fd(reqsock),
%     pass_fd(icmp6_fd), pass_fd(privproc_pair.second),
%     "teredo.remlab.net", "teredo.remlab.net"
% ])
% \end{lstlisting}

% In Listing \ref{miredo},
% we show non-trivial code for launching a real application:
% the Miredo IPv6 tunneling software.
% We use a few helper functions in this listing to keep the attention focused on the interesting parts.

% Miredo is separated into two components, a privileged process which sets up network interfaces,
% and an unprivileged process which talks to the network.
% With minimal modifications to Miredo,
% we launch Miredo entirely unprivileged inside a user namespace and network namespace,
% with all resources created outside and explicitly passed in.

%% **** note to self: concepts introduced in this example
%% socketpair
%% several helper functions
%% **** note to self
%% maybe I should remove some of the unnecessary networking stuff?
%% maybe show the corresponding fork-based code?
%% except doing something with fork with the same error handling requires a bunch of crazy IPC
\section{Implementation}\label{implementation}
\subsection{Basics about rsyscall}
Our main need for implementing direct-style process creation
is a robust system for cross-process syscalls.
We provided this in the rsyscall project.
rsyscall is a toolkit for cross-process syscalls on Linux,
with several language-specific library implementations.

@@

Rewrite some existing system with rsyscall?
Write a shell?

Right okay so we can do some fork-style code, sure.
Show the IPC, not too complex, but show it.

We can do it in Python with rsyscall - have a hidden experimental secret fork thing.
Using write_all_bytes will be annoying but fine.
Ugh, but we'll have to decode it too.
Eh, no, we can wrap that.
Yeah then it's fine.
\section{Future work}\label{future_work}
\subsection{Other applications of rsyscall}
Most avenues of future work focus on rsyscall.
rsyscall was not developed solely for the purpose of this paper,
and it has many uses unrelated to direct-style process creation,
such as asynchronous system calls, exceptionless system calls\cite{flexsc}, cross-host operations, among others.
We are actively exploring such applications,
as well as broadening rsyscall's language support.
\subsection{Kernel support}
rsyscall's cross-process syscalls can be performed entirely in userspace,
which has substantial benefits for deployability.
Nevertheless, direct kernel support for creating a stub process and performing syscalls in the context of that process
may provide efficiency benefits, as well as reducing userspace-visible complexity.

>>>>

Fork-style (such as \texttt{fork}) and spawn-style (such as \texttt{posix\_spawn}) each have advantages and disadvantages.
Correct and performant usage of fork-style places significant constraints on the calling process.
Spawn-style lifts many of these constraints,
but is fundamentally limited in the kinds of processes it can create.
With both fork-style and spawn-style,
it is difficult to
branch based on results or errors encountered while initializing the new process;
fork-style requires IPC if it wishes to communicate errors back to the rest of the program,
while robust indication of errors in spawn-style requires substantial growth in the spawn interface.
Additionally, modern operating systems such as Linux have features
which are difficult or impossible to fully exploit through fork-style or spawn-style.

Direct-style is rare in general, but relatively common on academic operating systems.
Direct-style involves creating a new child process in a non-running state,
and then calling individual syscalls from the parent to initialize the child.
Direct-style does not constrain the calling process,
easily branches at any step of the initialization,
and can create any kind of process supported by the underlying system.

Our contribution is an implementation of direct-style process creation for Linux,
making use of our work on the rsyscall project.
The rsyscall project develops language-specific ``process-independent'' syscall libraries,
replacing libc for interaction with Linux,
allowing a program to perform cross-process syscalls.
rsyscall currently has best support for Python,
but can be ported to other languages;
the examples we show in this paper will be in Python,
but generalize easily to other languages.

In our implementation, a parent process calls the \texttt{clone} syscall
to create a child process which shares memory and other resources with the parent,
and which runs no user code.
The parent then uses \texttt{rsyscall}
to call arbitrary syscalls in the child process to initialize it;
for example, the parent might call \texttt{chdir} to change the child's current working directory.
After the process is fully initialized,
the parent can call \texttt{execve} in the child,
causing the process to start running an arbitrary executable
and detaching the child from the control of \texttt{rsyscall}.

A direct-style \texttt{clone} has all the advantages of direct-style process creation
over Linux's fork-style \texttt{fork} syscall and spawn-style \texttt{posix\_spawn} libc function.
Direct-style \texttt{clone} can be used correctly and performantly from any process.
Cross-process syscalls made with rsyscall individually report errors, as normal, and can be easily branched on.
Arbitrary processes involving complex namespaces and initialization can be created
using techniques that are impossible with \texttt{fork}.

We have used direct-style process creation on Linux to do TODO and TODO,
and we've benchmarked our implementation as TODO.
Direct-style \texttt{clone} has minimal performance cost in normal situations,
and can outperform \texttt{fork} in some cases.
We have deployed direct-style process creation in production internally at Two Sigma.

Our implementation is distributed as a feature of \texttt{rsyscall},
and is immediately deployable on recent Linux kernels.
rsyscall is open source and available from \url{https://github.com/catern/rsyscall}.

This paper is organized in several sections.
In section \ref{background}, we provide background on fork-style, spawn-style, and direct-style process creation,
and consider their advantages and disadvantages in more depth.
In section \ref{examples} we show examples of usage of direct-style process creation in Python on Linux,
including use of process features which are difficult to use on Linux with \texttt{fork} or \texttt{posix\_spawn},
and briefly examine the user API.
In section \ref{implementation}, we explore our implementation in depth,
and discuss several difficult process-creation related details of \texttt{rsyscall}'s implementation.
In section \ref{evaluation}, we evaluate, somehow (TODO).
In section \ref{future_work},
we list several directions for future work related to direct-style process creation and to \texttt{rsyscall}.
In section \ref{conclusion}, we briefly and optimistically conclude.

@@

\section{Background on the use of processes}
Why is it important to have a high-quality interface for creating processes?
Processes are already widely used;
most software is distributed as an executable which runs in a dedicated process.
This basic usage of processes can be performed with even a complex and inefficient process creation interface.
But processes have many uses beyond this simple and widespread one;
here we examine some more sophisticated applications of processes,
which benefit from a better process creation interface.
\subsection{Abstraction over resources}
In Unix, the mechanism of file descriptor inheritance
allows a process to be provided a resource by its creator,
while abstracting over the precise identity of that resource.[fn:ucspi]
For example, a process can be provided a file descriptor,
which can correspond to any file in a filesystem,
without the process being aware of what specific file it is accessing.
This is further enhanced by Unix's "everything is a file" design;
the passed file descriptor could also be a pipe, a network connection, or some other resource,
without the process knowing.
As another example,
a process can be provided a socket file descriptor on which it can call \texttt{accept} to receive connections,
without being aware of whether those connections come from the internet or from a local Unix socket.[fn:ucspi]
This abstraction mechanism is the basic principle of pipelines and redirection in the Unix shell,
but it is rarely used outside of the shell.
\subsection{Customization without explicit support}
In many systems,
it's possible to modify a process's view of nominally "global" resources.
In Unix-derived systems, this ability is most influentially provided in Plan 9[fn:plan9],
which allows each process to customize its view of the filesystem with private mounts and union directories[fn:plan9ns].
In Linux, these concepts were implemented as per-process namespaces[fn:linuxns].
Fundamentally,
this allows customizing a process's environment and therefore a program's behavior,
without having to write explicit support code for customization.
For example, Plan 9, unlike most other Unix-derived systems,
did not have a \texttt{PATH} environment variable which was searched by code in the process to find executables;
instead, each process was executed with a \texttt{/bin} directory at the root of the filesystem,
which was a union of many other directories,
and simply executed \texttt{/bin/foo} to run the program named \texttt{foo}.
In this way the set of executables provided to a process could be customized,
without any code to parse and handle \texttt{PATH} or any other executable-lookup-specific customization code.
\subsection{Sandboxing}
The basic isolation powers of processes are used to simplify application development:
it is beneficial to have a private virtual memory space when developing a stand-alone program.
But most systems have additional mechanisms of isolation between processes,
such as different privilege levels and access to global resources,
which can be used to provide a form of sandboxing.
For example, components which may exposed to hostile network requests
can be run in a separate process, at a lower privilege level than the main program;
in this way, even if an attacker gains control over that component,
the attacker will only have access to the lower level of privileges of that component,
rather than the full privileges of the main program.
\subsection{Capability-based security}
As a further development of process-based sandboxing,
the privileges of a process can be explicitly enumerated
in a capability-based security model.[fn:capsicum]
By using previously-mentioned resource passing mechanisms,
such as file descriptor inheritance or namespace manipulation,
and by disabling the process's access to global resources such as the shared filesystem,
we can enforce that all resources used by the process are passed at creation time.
\subsection{Non-shared-memory concurrency}
Processes run concurrently,
which allows exploiting parallelism in the hardware.
Since processes don't share memory,
they can provide a less complex parallel programming environment
than shared-memory thread-based approaches.
The most popular parallel programming environment in existence today is the Unix shell,
which obtains its parallelism by running multiple processes connected via pipes.
The Unix shell has a relatively constrained form of parallel processing,
but it's also possible to create more complex webs of parallel processes,
where, for example, one process might take multiple inputs over multiple pipes,
or produce multiple outputs.
\subsection{Conclusion}
These techniques, and more, are available through the process interface.
Most software would benefit from abstraction over resources, sandboxing, and parallelism.
Yet these features of processes are used only rarely.
There are multiple reasons for this,
but one of the primary reasons is the complexity of current process creation interfaces.

Many of these techniques are used today by specialized software and services.
Often, such software only allows use of one of these techniques;
for example, the Unix shell allows piping together process, but not namespacing them;
container systems allow sandboxing processes, but not piping them together.
By delegating these features to specific separate services,
we lose the ability to use them in combination.

By improving the process creation interface,
we can make it possible both for programs to directly manipulate processes to use these techniques,
and to use and share composable libraries which use these techniques.
We believe this potential justifies the investment of substantial effort
into improving the process creation interface.
\section{Overview and examples}\label{examples}
In most operating system APIs, including the ones available on Linux,
when a program makes a syscall, it implicitly operates on the current process.
To perform direct-style process creation on Linux,
we need a Linux API where we instead explicitly specify in which process we want to make a syscall.

@@

\subsection{Example: Create and exec}
\begin{lstlisting}[float,language=Python,label={basic},caption={Creating a new process and immediately execing}]
# Use the clone syscall in the local thread to create a new thread;
# we use a wrapper that supplies defaults for all arguments.
child = local.thread.clone()
# Call execv to run a different executable in the child thread;
# We pass the executable path as the first argument in the argument list, as is traditional.
child.execv(hello_path, [hello_path])
\end{lstlisting}
In Listing \ref{basic}, we simply create a new process under our control (a thread)
and immediately exec a binary with it.
As exceptions are used for error-handling in the Python API,
there is no need for error-checking code.
%% **** note to self: concepts introduced in this example
%% local.thread
%% clone
%% execve
\subsection{Example: Passing down FDs}
\begin{lstlisting}[float,language=Python,label={fds},caption={Passing down FDs}]
child = local.thread.clone()
sock = child.socket(AF.INET, SOCK.DGRAM)
# Bind the socket to a sockaddr_in;
# the sockaddr is written to memory accessible by the child with child.ptr
sock.bind(child.ptr(SockaddrIn(0, 0)))
sock.listen(10)
sock.disable_cloexec()
child.execv(executable_path, [executable_path, "--listening-socket", str(int(sock))])
\end{lstlisting}
In Listing \ref{fds}, we create a new thread,
then create a listening socket bound to a random port in that thread,
then call exec, 
passing down the socket by disabling cloexec and passing its file descriptor number as an argument to the new program.

File descriptors, here, are object oriented and have relevant syscalls as methods.
They make syscalls in the process they are created in by default;
we can create more objects referring to the same file descriptor from different processes
if we want to make the syscalls from another process.
%% **** note to self: concepts introduced in this example
%% disable_cloexec
%% socket creation
%% using int on sock
%% .ptr
\subsection{Example: Pipeline}
\begin{lstlisting}[float,language=Python,label={pipe},caption={Creating a pipeline}]
# create the pipe
pipe = local.thread.pipe()
child1 = local.thread.clone()
# inherit the write-end of the pipe to child1, and replace child1.stdout with it
child1_pipe_write = child1.inherit_fd(pipe.write)
child1_pipe_write.dup2(child1.stdout)
child1_proc = child1.execv(yes_path, [yes_path])
child2 = local.thread.clone()
# inherit the read-end of the pipe to child2, and replace child2.stdin with it
child2_pipe_read = child2.inherit_fd(pipe.read)
child2_pipe_read.dup2(child2.stdin)
child2_proc = child2.execv(head_path, [head_path, "-n", "15"])
\end{lstlisting}
In Listing \ref{pipe},
we do the same as the Unix shell pipeline "yes | head -n 15".
We create a pipe,
then create two threads,
connect them with a pipe,
and exec a different program in each thread.

@@

We bind to a privileged port on localhost inside the namespace,
and create one child to listen on that socket,
and another child to connect to it.
\subsection{Example: Starting a complex system}
\begin{lstlisting}[float,language=Python,label={miredo},caption={Starting a complex system}]
### create socket outside network namespace that Miredo will use for internet access
inet_sock = local.thread.socket(AF.INET, SOCK.DGRAM)
inet_sock.bind(local.thread.ptr(SockaddrIn(0, 0)))
# set some miscellaneous additional sockopts that Miredo wants
set_miredo_sockopts(local.thread, inet_sock)
### create main network namespace thread
ns_thread = local.thread.clone(CLONE.NEWNET|CLONE.NEWUSER)
### create in-network-namespace raw INET6 socket which Miredo will use to relay pings
icmp6_fd = ns_thread.socket(AF.INET6, SOCK.RAW, IPPROTO.ICMPV6)
### create in-network-namespace socket,
### which Miredo will use for unassociated Ifreq ioctls
reqsock = ns_thread.socket(AF.INET, SOCK.STREAM)
### create and set up the TUN interface
tun_fd, tun_index = make_tun(ns_thread, "miredo", reqsock)
### create socketpair which Miredo will use to communicate
### between privileged process and Teredo client
privproc_pair = ns_thread.socketpair(AF.UNIX, SOCK.STREAM)
### start up privileged process which manipulates the network setup in the namespace
privproc_thread = ns_thread.clone()
# We preserve NET_ADMIN capability over exec so that privproc can
# manipulate the TUN interface. add_to_ambient_caps is a helper function
# used because manipulating Linux ambient capabilities is fairly verbose
add_to_ambient_caps(privproc_thread, {CAP.NET_ADMIN})
# privproc expects to communicate with the main client over stdin and stdout
privproc_side = privproc_thread.inherit_fd(privproc_pair.first)
privproc_side.dup2(privproc_thread.stdin)
privproc_side.dup2(privproc_thread.stdout)
privproc_child = privproc_thread.execv(miredo_privproc_executable_path, [
    miredo_privproc_executable_path, str(tun_index)
])
### start up Miredo client process which communicates
### over the internet to implement the tunnel
# the client process doesn't need to be in the same network namespace,
# since it is passed all the resources it needs as fds at startup.
client_thread = ns_thread.clone(CLONE.NEWUSER|CLONE.NEWNET|CLONE.NEWNS|CLONE.NEWPID)
# lightly sandbox by unmounting everything except for the executable
# and its deps (known via package manager)
unmount_everything_except(client_thread, miredo_exec.run_client.executable_path)
# a helper function for preparing the fds that are passed as command line arguments
async def pass_fd(fd: FileDescriptor) -> str:
    client_thread.inherit_fd(fd).disable_cloexec()
    return str(int(fd))
client_child = client_thread.execv(miredo_client_executable_path, [
    miredo_client_executable_path,
    pass_fd(inet_sock), pass_fd(tun_fd), pass_fd(reqsock),
    pass_fd(icmp6_fd), pass_fd(privproc_pair.second),
    "teredo.remlab.net", "teredo.remlab.net"
])
\end{lstlisting}

In Listing \ref{miredo},
we show non-trivial code for launching a real application:
the Miredo IPv6 tunneling software.
We use a few helper functions in this listing to keep the attention focused on the interesting parts.

Miredo is separated into two components, a privileged process which sets up network interfaces,
and an unprivileged process which talks to the network.
With minimal modifications to Miredo,
we launch Miredo entirely unprivileged inside a user namespace and network namespace,
with all resources created outside and explicitly passed in.
\section{Implementation}\label{implementation}
\subsection{Basics about rsyscall}
Our main need for implementing direct-style process creation
is a robust system for cross-process syscalls.
We provided this in the rsyscall project.
rsyscall is a toolkit for cross-process syscalls on Linux,
with several language-specific library implementations.

@@

Rewrite some existing system with rsyscall?
Write a shell?
\section{Future work}\label{future_work}
\subsection{Other applications of rsyscall}
Most avenues of future work focus on rsyscall.
rsyscall was not developed solely for the purpose of this paper,
and it has many uses unrelated to direct-style process creation,
such as asynchronous system calls, exceptionless system calls\cite{flexsc}, cross-host operations, among others.
We are actively exploring such applications,
as well as broadening rsyscall's language support.
\subsection{Kernel support}
rsyscall's cross-process syscalls can be performed entirely in userspace,
which has substantial benefits for deployability.
Nevertheless, direct kernel support for creating a stub process and performing syscalls in the context of that process
may provide efficiency benefits, as well as reducing userspace-visible complexity.
